---
title: "EpiTxDb.Hs.hg38: Annotation package for EpiTxDb objects"
author: "Felix G.M. Ernst"
date: "`r Sys.Date()`"
package: EpiTxDb.Hs.hg38
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    df_print: paged
vignette: >
  %\VignetteIndexEntry{EpiTxDb.Hs.hg38}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown(css.files = c('custom.css'))
```

# Available resources

`EpiTxDb.Hs.hg38` contains post-transcriptional RNA modifications from RMBase
v2.0 [[@Xuan.2017]](#References), tRNAdb [[@Juehling.2009]](#References) and
snoRNAdb [[@Lestrade.2006]](#References) and can be accessed through the 
functions `EpiTxDb.Hs.hg38.tRNAdb()`, `EpiTxDb.Hs.hg38.snoRNAdb()` and 
`EpiTxDb.Hs.hg38.RMBase()`

```{r, echo=FALSE}
suppressPackageStartupMessages({
  library(EpiTxDb.Hs.hg38)
})
```
```{r, eval=FALSE}
library(EpiTxDb.Hs.hg38)
```
```{r}
etdb <- EpiTxDb.Hs.hg38.tRNAdb()
etdb
```

Modification information can be accessed through the typical function for an
`EpiTxDb` object, for example `modifications()`:

```{r}
modifications(etdb)
```

For a more detailed overview and explanation of the functionality of the 
`EpiTxDb` class, have a look at the `EpiTxDb` package.

# Chain files for rRNA

The rRNA sequence annotation for ribosomal RNA has undergone some clarification
processes in recent years. Therefore some of the annotation of modification 
refer to an older rRNA annotation.

In order to help using and updating older information, a chain file was
generated for use with `liftOver`, which allows conversion of hg19 coordinates
to hg38 coordinates and back. The resources can be loaded via
`chain.rRNA.hg19Tohg38()` and `chain.rRNA.hg38Tohg19()`.

```{r}
cf <- chain.rRNA.hg19Tohg38()
cf
```

The following example illustrate a use case, in which data from the Modomics
[[@Boccaletto.2018]](#References) database will be used. The sequence data
currently stored, is the hg19 version. First we load the sequence as
`ModRNAStringSet`.

```{r}
library(rtracklayer)
library(Modstrings)
files <- c(system.file("extdata","Modomics.LSU.Hs.txt",
                       package = "EpiTxDb.Hs.hg38"),
           system.file("extdata","Modomics.SSU.Hs.txt",
                       package = "EpiTxDb.Hs.hg38"))
seq <- lapply(files,readLines,encoding = "UTF-8")
seq <- unlist(seq)
names <- seq[seq.int(1L,6L,2L)]
seq <- seq[seq.int(2L,6L,2L)]
seq <- ModRNAStringSet(sanitizeFromModomics(gsub("-","",seq)))
names(seq) <- c("28S","5.8S","18S")
mod <- separate(seq)
```

The position for the one `m7G` and two `m6A` are of for the current rRNA
sequences. This is also the case for the other modifications.

```{r}
mod[mod$mod == "m7G" | mod$mod == "m6A"]
```

Using the chain file we can `liftOver` the coordinates, which matches the 
expected coordinates.

```{r}
mod_new <- unlist(liftOver(mod,cf))
mod_new[mod_new$mod == "m7G" | mod_new$mod == "m6A"]
```

In addition, the `ModRNAStringSet` object can be update with the current 
sequence.

```{r}
rna <- snoRNA.targets.hg38()
names(rna)[1:4] <- c("5S","18S","5.8S","28S")
seqtype(rna) <- "RNA"
seq_new <- combineIntoModstrings(rna, mod_new)
seq_new
```

# Sessioninfo

```{r}
sessionInfo()
```

<a name="References"></a>

# References
